### 2014-07-24 Pero_Transcriptome.tex
mystery update

% Title must be 150 characters or less
\begin{flushleft}
{\Large
\textbf{Characterization of the transcriptome, nucleotide sequence polymorphism and selection in the desert adapted mouse \textit{Peromyscus eremicus}}
}
% Insert Author names, affiliations and corresponding author email.
\\
Matthew D. MacManes$^{1}$, 
Michael B. Eisen $^{2}$, 
\\
\bf{1} Department of Molecular, Cellular and Biomedical Sciences, University of New Hampshire Institution Name, Durham NH, USA
\\
\bf{2} HHMI and University of California, Berkeley, Berkeley, CA, USA
\\
$\ast$ E-mail: macmanes@gmail.com, @PeroMHC
\end{flushleft}

@@

For biologists interested in understanding the relationship between fitness, genotype, and phenotype, modern sequencing technologies provide for an unprecedented opportunity to gain a deep understanding of genome level processes that together, underlie adaptation. Unlike more traditional approaches

 One interesting example of adaptation lies in animals ability to survive desert conditions. Here, heat \emph{and} drought provide for powerful selective forces, testing animals' ability to osmoregulate and thus to survive. \\

Specifically, the maintenance of water balance in animals is one of the most important physiologic processes, and is critical to desert survival \cite{Walsberg:2000uu}. Indeed, mammals are exquisitely sensitive to changes in osmolality, with slight derangement eliciting physiologic compromise.  When the loss of water exceeds dietary intake, dehydration - and in extreme cases, death - can occur.  Unlike most mammals, animals living in desert habitats are subjected to long periods of extreme heat and intense drought.  As a result, desert animals have evolved mechanisms through which physiologic homeostasis is maintained despite severe and prolonged dehydration. \\

One such desert-adapted rodent, a cricetid rodent endemic to the Southwest United States \cite{Veal:2001vp}, is a novel model for the study of adaptation to desert environments. They have a lifespan typical of small mammals, and therefore an individual may live it's entire life without ever drinking water. These rodents have distinct advantage over other desert animals (e.g. \textit{Dipodomys}) in that they breed readily in captivity, which enables laboratory studies of the phenotype of interest. In addition the focal species is positioned in a clade of well known animals (e.g. \textit{P. californicus}, \textit{P. maniculatus} and \textit{P. polionotus}) \cite{Bradley:2007jo}. There are growing genetic and genomic resources available \cite{Shorter:2014kw,Panhuis:2011co,Shorter:2012ib}. \\

>>>>

% Title must be 150 characters or less
\begin{flushleft}
{\Large
\textbf{Characterization of the transcriptome, nucleotide sequence polymorphism, and natural selection in the desert adapted mouse \textit{Peromyscus eremicus}}
}
% Insert Author names, affiliations and corresponding author email.
\\
Matthew D. MacManes$^{1}$, 
Michael B. Eisen $^{2}$, 
\\
\bf{1} Department of Molecular, Cellular and Biomedical Sciences, University of New Hampshire, Durham NH, USA
\\
\bf{2} HHMI and University of California, Berkeley, Berkeley, CA, USA
\\
$\ast$ E-mail: macmanes@gmail.com, @PeroMHC
\end{flushleft}

@@

For biologists interested in understanding the relationship between fitness, genotype, and phenotype, modern sequencing technologies provide for an unprecedented opportunity to gain a deep understanding of genome level processes that together, underlie adaptation. Unlike more traditional approaches (e.g. QTL mapping), second generation sequencing allow researchers to access genome wide patterns of gene expression, nucleotide variation and ulitimately pattens of natural selection in non-model organisms lacking an extensive set of genomic tools. This approach has been used widely in recent years in characterizing X, Y and Z. \\

One interesting example of adaptation lies in animal's ability to survive desert conditions. Here, heat \emph{and} drought provide for powerful selective forces, testing animals' ability to osmoregulate and thus to survive \cite{Walsberg:2000uu}. Indeed, the maintenance of water balance in animals is one of the most important physiologic processes, and is critical to desert survival. Mammals are exquisitely sensitive to changes in osmolality, with slight derangement eliciting physiologic compromise.  When the loss of water exceeds dietary intake, dehydration - and in extreme cases, death - can occur.  Unlike most mammals, animals living in desert habitats are subjected to long periods of extreme heat and intense drought.  As a result, desert animals have evolved mechanisms through which physiologic homeostasis is maintained despite severe and prolonged dehydration. Understanding these mechanismsm will significantly enhance our understanding of the physiologic processes underlying osmoregulation in extreme environments, having implications for studies of human health, conservation, and climate change. \\

Genes responsible for the maintinance of water balance are well characterized in model organisms such as mice \cite{Tatum:2009kp}, rats \cite{Romero:2007fs,Rojek:2006dg,NIELSEN:1995uq}, and humans \cite{Mobasheri:2007tt,Bedford:2003wa,NIELSEN:1999vg}. In addition to these studies, a ong standing interest in deserts adaptation has resulted in a number of studies that looked at the morphology or expression of single genes in the renal tissues of desert adapted rodents \textit{Phyllotis darwini} \cite{Gallardo:2005fm}, \textit{Psammomys obesus} \cite{Kaissling:1975uu}, and \textit{Perognathus penicillatus} \cite{Altschuler:1979du}. More recently, full renal transcriptomes have been generated for \textit{Dipodomys spectabilis} and \textit{Chaetodipus baileyi} \cite{Marra:2014de} as well as \textit{Abrothrix olivacea} \cite{Giorello:2014iv}. \\

These studies provide a rich context for the current and future work, aimed at developing a synthetic understanding of the the genetic and genomic underpinnings of desert adaptation in rodents. As a first step, we have sequenced, assembled, and characterized the transcriptome (using four tissue types - liver, kidney, testes, brain), of a desert adapted cricetid rodent endemic to the Southwest United States \cite{Veal:2001vp}, \textit{Peromyscus eremicus}. These animals have a lifespan typical of small mammals, and therefore an individual may live it's entire life without ever drinking water. These rodents have distinct advantage over other desert animals (e.g. \textit{Dipodomys}) in that they breed readily in captivity, which enables future laboratory studies of the phenotype of interest. In addition the focal species is positioned in a clade of well known animals (e.g. \textit{P. californicus}, \textit{P. maniculatus} and \textit{P. polionotus}) \cite{Bradley:2007jo} with growing genetic and genomic resources \cite{Shorter:2014kw,Panhuis:2011co,Shorter:2012ib} which together suggest that future comparative studies are possible. \\
